A Unified Treatment of Tractability for \\ Approximation Problems Defined on Hilbert Spaces

### 2024-04-21 reply_reviewers.tex

\begin{document}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\begin{center}
    {\Large Response to the reviewers of the manuscript\\``A Unified Treatment of Tractability for \\ Approximation Problems Defined on Hilbert Spaces''}
\end{center}

We would like to thank the referees for their careful review and their valuable suggestions. Following the advice of the referees, we give an overview of the changes made in the manuscript.

@@

\subsection*{Referee Report 1 (1 page):}
\begin{itemize}
    \item We are not quite sure whether we understood this remark correctly, but we have tried to reformulate the corresponding sentence in the introduction. 
    \item We assume the referee means p.~2. We have re-formulated this sentence. 
    \item We assume the referee means p.~2. We have added a footnote with an explanation. 
    \item We have added another appendix (now, Appendix 1), where we show an example of how the condition for strong exponential tractability derived in our paper (which is different to what had been known before) can be applied to a concrete problem setting. The proof of Proposition 1 has been moved to Appendix 2. 
\end{itemize}
\medskip

\subsection*{Referee Report 2 (3 pages):}
\begin{itemize}
    \item[1.] We have shortened the paragraph. 
    \item[2.] We have added explanations to the introductory parts of the paper, and we have followed the suggestion of the referee to include a simplified version of the table for illustration. 
    \item[3.] We have re-formulated the sentence that was on p.~2, line~25. 
    \item[4.] A statement has been added as regards the role of $\vp$.
    \item[5.] Page 3, towards the top an explanation has been added.
    \item[6.] Page 2 (footnote) we explain that for problems where $\SOL_d$ has large singular values, a large error tolerance may be a reasonable expectation.  Thus we do not want to put an upper bound on $\varepsilon$.
    \item[7.] We have added a corresponding statement on page 3. 
    \item[8.] We have added a comment on this. 
    \item[9.] We have added a corresponding statement. 
    \item[10.] Done. 
    \item[11.] The referee is right. We have made the corresponding corrections. 
    \item[12.] We have removed the statement. 
    \item[13.] We added the precise equation we are using.
    \item[14.] We would like to emphasize that we intend to utilize equation (41) prior to establishing $\thed$ = 1. Indeed, we will make use of (41) in its general form in the proof of Theorem 5. 
    \item[15.] Again, the referee is right. We have made the formulation more precise. 
    \item[16.] Page 16, line -9: Done.
    \item[17.] This typo has been corrected. That is, we have changed ``Equation (39)'' to ``Equation (45)''.
    \item[18.] Done.
\end{itemize}

\bigskip

Apart from these changes, we removed a few typos and minor errors from the manuscript.

\medskip

We hope that these modifications address the concerns of the referees in a satisfactory manner. 
All relevant changes are marked in blue in the revised version.\\

\noindent With kind regards,\\[0.25cm]

\noindent O.~Emenike, F.J.~Hickernell, P.~Kritzer

\end{document}

### 2024-04-21 GeneralizedTractability_rev_2.tex
Updates from Overleaf

\newcommand{\bsa}{\boldsymbol{a}}
\newcommand{\bsb}{\boldsymbol{b}}
\newcommand{\bsh}{\boldsymbol{h}}
\newcommand{\kor}{{\mathrm{kor}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
%\newcommand{\norm}[1]{\left\Vert#1\right\Vert}

\input{FJHDef.tex}

@@

\newcommand{\fred}[1]{\begingroup\color{blue}#1\endgroup}
\newcommand{\peter}[1]{\begingroup\color{violet}#1\endgroup}
\newcommand{\kachi}[1]{\begingroup\color{ForestGreen}#1\endgroup}

@@

\fred{The authors have identified common themes in the proofs of necessary and sufficient conditions for tractability that appear in the literature.  By abstracting these themes, this article aims to unify and generalize many existing tractability results for approximation problems where all possible linear functional data are available. This is done in a series of five theorems, which provide necessary and sufficient conditions for (strong) tractability.}

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and let $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation*} %\label{eq:eigenvalues}
 \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation*}
\fred{Here, $\naturals$ denotes the set of positive integers.}  This means that \peter{the singular values of the operator $\SOL_d$ are the non-negative $\lambda_{i,d}$}. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.

@@

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$, defined on the unit ball of functions in $\cf_d$, denoted $\cb_d$, which  satisfy an absolute error criterion:
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty)\footnote{We note that the radius of the $\cb_d$ could be taken to be an arbitrary positive value, $R$, in which case the problem becomes equivalent to the original one with the tolerance $\varepsilon$ replaced by $\varepsilon/R$ since the problem and the optimal approximate solution are both linear (see below). \fred{We allow the error tolerance, $\varepsilon$, to be arbitrarily large because if $\SOL_d$ has large singular values, a large error tolerance may be a reasonable expectation.}}.
\end{equation}
Here, $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

The information complexity of the problem $\SOL_d$ is given by the function 
\[
\comp:(0,\infty) \times \naturals \to \natzero, 
\]
where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best admissible algorithm to satisfy the error criterion \eqref{eq:errcrit}\fred{, and  $\naturals_0$ denotes the set of non-negative integers}.
By definition, the information complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity since $d$ is typically the number of variables for the functions in $\cf_d$.

@@

\peter{Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$ (see \cite{NovWoz08a} for details). For the sake of brevity, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting (all criteria would then be normalized by the first singular value $\lambda_{1,d}$).}

@@

\peter{We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$, \kachi{$d$, and some parameter $\vp$}, and then identify conditions on the singular values that are equivalent to the fact that this bound holds, and are easier to verify than showing the bound directly. By ``simple'' we mean that $T$ takes a form such that the conditions in the theorems of this paper can be checked effectively; however, we remark that this is also related to the singular values $\lambda_{i,d}$ of the solution operator. Depending on those, a function $T$ may be useful for one particular problem setting but not for another.} \kachi{In Table \ref{tab:sampleT}, we present various common notions of tractability studied in the literature, characterized by $T$.} 
\begin{table}[hbt!]
    \caption{\kachi{Common notions of tractability}}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		\\
		\toprule
		\text{algebraic strong polynomial} & \max\{1,\varepsilon^{-p}\} \\
		\text{algebraic polynomial} & \max\{1,\varepsilon^{-p}\}d^{q} \\
		\text{exponential strong polynomial} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p \\
		\text{exponential polynomial} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} \\
        \text{algebraic quasi-polynomial} &
        \exp\{p(1+\log(\max\{1, \varepsilon^{-1}\}))(1+\log(d))\} \\
        \text{exponential quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} 
	\end{array}
\end{equation*}}
\label{tab:sampleT}
\end{table}

@@

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant\footnote{\peter{By ``constant'' we mean that the term is independent of the \emph{variables} $\varepsilon$ and $d$, but may depend on the \emph{parameter} $\vp$.}}, $C_{\vp}$, depending only on the parameter $\vp$. \kachi{For instance, when considering (algebraic) polynomial tractability, the parameter $\vp$ represents the exponents $(p,q)$ of $\varepsilon$ and $d$, as illustrated in Table 1}.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$. \fred{We write $\vzero$ to denote a vector with all components equal to zero, and $\vinfty$ to denote a vector with all components equal to $\infty$.
Furthermore, for two vectors $\vp=(p_1,\ldots,p_s)$ and $ \vp'=(p_1',\ldots,p_s')$ in $[0,\infty)^s$, 
we write $\vp\ge \vp'$ if $p_j\ge p'_j$ for all $j\in \{1,\ldots,s\}$. Furthermore, $\vp > \vp'$ 
if $p_j> p'_j$ for all $j\in \{1,\ldots,s\}$. The expression $\vp\in (\vp',\vp'')$ is to be interpreted as 
$\vp' < \vp < \vp''$ in that sense, and analogously for half-closed or closed intervals.}

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} below all take the following form:  for a given $\thed \in \{1,d\}$ and $\Omega \subseteq (0,\infty) \times \mathbb{N}$,
\begin{multline} \label{eq:metatheorem}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} >0 \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \text{where } \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0,
        \end{cases}
\end{multline}
%The differences in these theorems are in the choices of $\thed$ and $\Omega$. 
\peter{where $T(0,\thed,\vp)$ is defined as in \eqref{eq:T0} below.}  \fred{These theorems say essentially that \emph{(strong) tractability is equivalent to summability conditions on the singular values.} A more stringent form of tractability resulting from a smaller $T$ is equivalent to a more stringent summability condition since the terms being summed become larger.}

Although some of \Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} are special cases of others, we prove each separately to allow the reader to become familiar with the arguments used as additional layers of complexity are added. \peter{(This may be, e.g., additional
dependence on $d$ when going from strong tractability to tractability in Subsection~\ref{sec:tractability}, or a restriction of the domain of $\varepsilon$ for which we define tractability in Section 
~\ref{sec:restricted}).}  \Cref{thm:subhT} has a somewhat different form than \eqref{eq:metatheorem} above.

\fred{Although in many cases $\Omega$ is chosen to be all of $(0,\infty) \times \mathbb{N}$, there are situations where it makes sense to choose $\Omega$ as a proper subset of all possible values of $\varepsilon^{-1}$ and $d$.  For some problems with unbounded $d$---as in finance---there may be no need for arbitrarily small error tolerances, so $\Omega = (0,\varepsilon_{\min}^{-1}] \times \mathbb{N}$ may be appropriate.  For other problems with a moderate upper bound on $d$, $\Omega = (0,\infty) \times \{1, \ldots, d_{\max}\}$ may be appropriate. }

The remainder of this paper is organized as follows. In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when there is an upper bound on the information complexity that is independent of the dimension, $d$. We then turn to tractability  when the bound on the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and introduce other notions of tractability. A generalized notion of weak tractability is given in Section \ref{sec:subh}. In Section \ref{sec:restricted}, we consider tractability on a restricted domain, $\Omega$. \peter{There are two appendices to the paper. Appendix 1 contains an example that shows how we can make use of one of the findings in this paper in a concrete example. Appendix 2 contains the technical proof of Proposition \ref{prop:equiv_conditions}}.

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing $\varepsilon$, the tolerance, or increasing $d$, the dimension\footnote{This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.}. Furthermore, increasing $\vp$ allows for a possibly looser bound on the information complexity. Since there are an infinite number of positive singular values, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis in \Cref{sec:spt,sec:restricted}.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp)   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}
\peter{We remark that the latter assumption is quite natural, and is, for example fullfilled for all tractability types in Table~\ref{tab:commonT}.}

@@

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.  Fix any $\varepsilon > 0$ and any $d \in \naturals$, and let $\thed \in \naturals$ be arbitrary. %\delta replaced d'
Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)},
\end{align*}
\peter{where the infinite sum is guaranteed to exist by the condition \eqref{eq:strong_tractiff3}.}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, N \in \naturals\ \peter{\mbox{and}\ n\ge N}.
\end{multline}
%We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer and $n - N +1$ is assumed to be positive.

@@

The following proposition states that the conditions \eqref{eq:exp_spt} and \eqref{eq:KW_exp_spt}, as well as the conditions \eqref{eq:exp_pt} and \eqref{eq:KW_exp_pt}, are indeed equivalent. 
The proof of this result is deferred to \peter{Appendix 2} where we will prove this statement for the latter two conditions, i.e., for the conditions on exponential polynomial tractability. The case of exponential strong polynomial tractability is not explicitly included, as it is actually simpler and can be treated analogously.  \peter{We give a short example of how we can apply the condition \eqref{eq:exp_spt} to a concrete problem in Appendix 1.}

>>>>

\newcommand{\bsa}{\boldsymbol{a}}
\newcommand{\bsb}{\boldsymbol{b}}
\newcommand{\bsh}{\boldsymbol{h}}
\newcommand{\bsx}{\boldsymbol{x}}
\newcommand{\kor}{{\mathrm{kor}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
%\newcommand{\norm}[1]{\left\Vert#1\right\Vert}

\input{FJHDef.tex}
% Modify the elsarticle-num-names bibliography style to sort references alphabetically
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother

@@

\newcommand{\fred}[1]{\begingroup\color{blue}#1\endgroup}
\newcommand{\peter}[1]{\begingroup\color{violet}#1\endgroup}
\newcommand{\kachi}[1]
{\begingroup\color{ForestGreen}#1\endgroup}
\newcommand{\review}[1]{\begingroup\color{blue}#1\endgroup}

@@

The authors have identified common themes in the proofs of necessary and sufficient conditions for tractability that appear in the literature.  By abstracting these themes, this article aims to unify and generalize many existing tractability results for approximation problems where all possible linear functional data are available. This is done in a series of five theorems, which provide necessary and sufficient conditions for (strong) tractability.

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and let $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation*} %\label{eq:eigenvalues}
 \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation*}
Here, $\naturals$ denotes the set of positive integers.  This means that the singular values of the operator $\SOL_d$ are the non-negative $\lambda_{i,d}$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.

@@

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$, defined on the unit ball of functions in $\cf_d$, denoted $\cb_d$, which  satisfy an absolute error criterion:
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty).\footnote{We note that the radius of the $\cb_d$ could be taken to be an arbitrary positive value, $R$, in which case the problem becomes equivalent to the original one with the tolerance $\varepsilon$ replaced by $\varepsilon/R$ since the problem and the optimal approximate solution are both linear (see below). We allow the error tolerance, $\varepsilon$, to be arbitrarily large because if $\SOL_d$ has large singular values, a large error tolerance may be a reasonable expectation.}
\end{equation}
Here, $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

The information complexity of the problem $\SOL_d$ is given by the function 
\[
\comp:(0,\infty) \times \naturals \to \natzero, 
\]
where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best admissible algorithm to satisfy the error criterion \eqref{eq:errcrit}, and  $\naturals_0$ denotes the set of non-negative integers.
By definition, the information complexity is non-decreasing as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity since $d$ is typically the number of variables for the functions in $\cf_d$.

@@

Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$ (see \cite{NovWoz08a} for details). For the sake of brevity, we restrict ourselves to the absolute error criterion here, but presumably, analogous results hold for the normalized setting (all criteria would then be normalized by the first singular value $\lambda_{1,d}$).

@@

We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$, $d$, and some parameter $\vp$, and then identify conditions on the singular values that are equivalent to the fact that this bound holds, and are easier to verify than showing the bound directly. By ``simple'' we mean that $T$ takes a form such that the conditions in the theorems of this paper can be checked effectively; however, we remark that this is also related to the singular values $\lambda_{i,d}$ of the solution operator. Depending on those, a function $T$ may be useful for one particular problem setting but not for another. In Table \ref{tab:sampleT}, we present various common notions of tractability studied in the literature, characterized by $T$. 
\begin{table}[hbt!]
    \caption{Common notions of tractability}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		\\
		\toprule
		\text{algebraic strong polynomial} & \max\{1,\varepsilon^{-p}\} \\
		\text{algebraic polynomial} & \max\{1,\varepsilon^{-p}\}d^{q} \\
		\text{exponential strong polynomial} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p \\
		\text{exponential polynomial} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} \\
        \text{algebraic quasi-polynomial} &
        \exp\{p(1+\log(\max\{1, \varepsilon^{-1}\}))(1+\log(d))\} \\
        \text{exponential quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} 
	\end{array}
\end{equation*}}
\label{tab:sampleT}
\end{table}

@@

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant\footnote{By ``constant'' we mean that the term is independent of the \emph{variables} $\varepsilon$ and $d$, but may depend on the \emph{parameter} $\vp$.}, $C_{\vp}$, depending only on the parameter $\vp$. For instance, when considering (algebraic) polynomial tractability, the parameter $\vp$ represents the exponents $(p,q)$ of $\varepsilon$ and $d$, as illustrated in Table \ref{tab:sampleT}.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$. We write $\vzero$ to denote a vector with all components equal to zero, and $\vinfty$ to denote a vector with all components equal to $\infty$.
Furthermore, for two vectors $\vp=(p_1,\ldots,p_s)$ and $ \vp'=(p_1',\ldots,p_s')$ in $[0,\infty)^s$, 
we write $\vp\ge \vp'$ if $p_j\ge p'_j$ for all $j\in \{1,\ldots,s\}$. Furthermore, $\vp > \vp'$ 
if $p_j> p'_j$ for all $j\in \{1,\ldots,s\}$. The expression $\vp\in (\vp',\vp'')$ is to be interpreted as 
$\vp' < \vp < \vp''$ in that sense, and analogously for half-closed or closed intervals.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} below all take the following form:  for a given $\thed \in \{1,d\}$ and $\Omega \subseteq (0,\infty) \times \mathbb{N}$,
\begin{multline} \label{eq:metatheorem}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} >0 \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \text{where } \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0,
        \end{cases}
\end{multline}
%The differences in these theorems are in the choices of $\thed$ and $\Omega$. 
where $T(0,\thed,\vp)$ is defined as in \eqref{eq:T0} below. These theorems say essentially that \emph{(strong) tractability is equivalent to summability conditions on the singular values.} A more stringent form of tractability resulting from a smaller $T$ is equivalent to a more stringent summability condition since the terms being summed become larger.

Although some of \Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} are special cases of others, we prove each separately to allow the reader to become familiar with the arguments used as additional layers of complexity are added. (This may be, e.g., additional
dependence on $d$ when going from strong tractability to tractability in Subsection~\ref{sec:tractability}, or a restriction of the domain of $\varepsilon$ for which we define tractability in Section 
~\ref{sec:restricted}).  \Cref{thm:subhT} has a somewhat different form than \eqref{eq:metatheorem} above.

Although in many cases $\Omega$ is chosen to be all of $(0,\infty) \times \mathbb{N}$, there are situations where it makes sense to choose $\Omega$ as a proper subset of all possible values of $\varepsilon^{-1}$ and $d$.  For some problems with unbounded $d$---as in finance---there may be no need for arbitrarily small error tolerances, so $\Omega = (0,\varepsilon_{\min}^{-1}] \times \mathbb{N}$ may be appropriate.  For other problems with a moderate upper bound on $d$, $\Omega = (0,\infty) \times \{1, \ldots, d_{\max}\}$ may be appropriate.

The remainder of this paper is organized as follows. In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when there is an upper bound on the information complexity that is independent of the dimension, $d$. We then turn to tractability when the bound on the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and introduce other notions of tractability. A generalized notion of weak tractability is given in Section \ref{sec:subh}. In Section \ref{sec:restricted}, we consider tractability on a restricted domain, $\Omega$. There are two appendices to the paper. Appendix 1 contains an example that shows how we can make use of one of the findings in this paper in a concrete example. Appendix 2 contains the technical proof of Proposition \ref{prop:equiv_conditions}.

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing $\varepsilon$, the tolerance, or increasing $d$, the dimension\footnote{This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.}. Furthermore, increasing $\vp$ allows for a possibly looser bound on the information complexity. Since there are an infinite number of positive singular values, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis in \Cref{sec:spt,sec:restricted}.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp)   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}
We remark that the latter assumption is quite natural, and is, for example fulfilled for all tractability types in Table~\ref{tab:commonT}.

@@

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.  Fix any $\varepsilon > 0$ and any $d \in \naturals$, and let $\thed \in \naturals$ be arbitrary. %\delta replaced d'
Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)},
\end{align*}
where the infinite sum is guaranteed to exist by the condition \eqref{eq:strong_tractiff3}.
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, N \in \naturals\ \mbox{and}\ n\ge N.
\end{multline}
%We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer and $n - N +1$ is assumed to be positive.

@@

The following proposition states that the conditions \eqref{eq:exp_spt} and \eqref{eq:KW_exp_spt}, as well as the conditions \eqref{eq:exp_pt} and \eqref{eq:KW_exp_pt}, are indeed equivalent. 
The proof of this result is deferred to Appendix 2 where we will prove this statement for the latter two conditions, i.e., for the conditions on exponential polynomial tractability. The case of exponential strong polynomial tractability is not explicitly included, as it is actually simpler and can be treated analogously.  We give a short example of how we can apply the condition \eqref{eq:exp_spt} to a concrete problem in Appendix 1.
